---
title: "TimeSeries"
author: "Alejandro Aísa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

The following code corresponds to the first part of the section I of the masters' thesis. The visualization of the number of protests in the region of Northern Ireland in different formats. 

### Libraries 

```{r}
library(tidyverse)
library(readr)
library(lubridate)
require(tidyquant)
```

### Data 

As usual, we may load all the data stored in the repository. As stated in the previous step some specific featuring will be performed at each step. In this case, the number of protests per month in various contexts are to be calculated. 

#### Monthly protests - District aggregated 

```{r}
NI <- read_csv("Databases/Final_Derry.csv") %>% 
  select(-...1)  %>% 
  group_by(month = floor_date(week, unit = "month"), District) %>% 
  summarise(
    monthly_protests = sum(n_protests)
  ) %>% 
  ungroup()
```

#### Montly protests - NI

```{r}
total <- NI %>% group_by(month) %>% 
  summarise(
    total_protests = sum(monthly_protests)
  )
```

#### Derry 

```{r}
Derry <- NI %>% 
  filter(District == "Derry") 
```

#### Belfast 

```{r}
Belfast <- NI %>% 
  filter(District == "Belfast")
```

#### RDD

```{r}
RDD <- total %>% 
  filter(month >= "1983-01-01" & month <= "1990-12-31")
```

### Time Series 

```{r}
sysfonts::font_add_google("Playfair Display", family = "playfair display")
sysfonts::font_add_google("Gideon Roman", family = "gideon roman")
showtext::showtext_auto()  
```

#### Montly protests 

First, the time series depicting the number of protests per month in the region of Northern Ireland. 

```{r}
pi <- ggplot(data = total, aes(x = month, y = total_protests )) +
  geom_line()+
  geom_vline(xintercept = as.numeric(Derry$month[86]), linetype = 4)+
  geom_ma(n = 6, color = "red")+
  theme_classic()+
  labs(
    title = "Figure 1: Protests in Northern Ireland",
    subtitle = 
      "Evolution in time of the monthly number of protests in Northern Ireland",
    x = "year", 
    y = "number of protests", 
    caption = "Source: Own work and Ron Francisco's database")+
  theme(
    plot.title = element_text(face = "bold", 
                              size = 14, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    plot.subtitle = element_text(size = 12, 
                                 hjust = 0.5, 
                                 family = "gideon roman"),
    plot.caption = element_text(size = 8, family = "gideon roman"),
    axis.title.x = element_text(size = 11, family = "gideon roman"), 
    axis.title.y = element_text(size = 11, family = "gideon roman"))
```


```{r}
pi
```
#### Montly protests in Derry 

Second time series: number of protests per month in the Derry District. 

```{r}
pd <- ggplot(data = Derry, aes(x = month, y = monthly_protests )) +
  geom_line()+
  geom_vline(xintercept = as.numeric(Derry$month[86]), linetype = 4)+
  geom_ma(n = 6, color = "red")+
  theme_classic()+
  labs(
    title = "Figure 2: Protests in Derry",
    subtitle = 
      "Evolution in time of the monthly number of protests in Derry",
    x = "year", 
    y = "number of protests", 
    caption = "Source: Own work and Ron Francisco's database")+
   theme(
    plot.title = element_text(face = "bold", 
                              size = 14, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    plot.subtitle = element_text(size = 12, 
                                 hjust = 0.5, 
                                 family = "gideon roman"),
    plot.caption = element_text(size = 8, family = "gideon roman"),
    axis.title.x = element_text(size = 11, family = "gideon roman"), 
    axis.title.y = element_text(size = 11, family = "gideon roman"))

pd
```

#### Mothly protests in Belfast 

Third visual representation: Number of protests per month in the District of Belfast. 

```{r}
pb <- ggplot(data = Belfast, aes(x = month, y = monthly_protests )) +
  geom_line()+
  geom_vline(xintercept = as.numeric(Derry$month[86]), linetype = 4)+
  geom_ma(n = 6, color = "red")+
  theme_classic()+
  labs(
    title = "Figure 3: Protests in Belfast",
    subtitle = 
      "Evolution in time of the monthly number of protests in Belfast",
    x = "year", 
    y = "number of protests", 
    caption = "Source: Own work and Ron Francisco's database")+
   theme(
    plot.title = element_text(face = "bold", 
                              size = 14, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    plot.subtitle = element_text(size = 12, 
                                 hjust = 0.5, 
                                 family = "gideon roman"),
    plot.caption = element_text(size = 8, family = "gideon roman"),
    axis.title.x = element_text(size = 11, family = "gideon roman"), 
    axis.title.y = element_text(size = 11, family = "gideon roman"))

pb
```

#### Discontinuity (Annex I)

Additional representation accounting for the regression discontinuity design. The monthly number of protests in NI in the period 1983-1991 are present. 

```{r}
prdd <- ggplot(data = RDD, aes(x = month, y = total_protests )) +
  geom_line()+
  geom_vline(xintercept = as.numeric(Derry$month[86]), linetype = 4)+
  geom_ma(n = 6, color = "red")+
  theme_classic()+
  labs(
    title = "Figure 4: Protests before and after Derry FC",
    subtitle = 
      "Evolution in time of the monthly number of protests in Northern Ireland",
    x = "year", 
    y = "number of protests", 
    caption = "Source: Own work and Ron Francisco's database")+
   theme(
    plot.title = element_text(face = "bold", 
                              size = 14, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    plot.subtitle = element_text(size = 12, 
                                 hjust = 0.5, 
                                 family = "gideon roman"),
    plot.caption = element_text(size = 8, family = "gideon roman"),
    axis.title.x = element_text(size = 11, family = "gideon roman"), 
    axis.title.y = element_text(size = 11, family = "gideon roman"))

prdd
```

