---
title: "MapViz"
author: "Alejandro Aísa"
date: "`r Sys.Date()`"
output: pdf_document
---

The following code corresponds to the second part of the section I of the masters' thesis. The geographical representation of the number of protests in the region of Northern Ireland in different maps. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(sp)
library(rgdal)
library(rgeos)
library(sf)
library(lubridate)
require(gganimate)
require(transformr)
require(ggthemes)
library(ggspatial)
library(tidyverse)
```

## Data 

As usual, we may load all the data stored in the repository. As stated before,  some specific featuring will be performed at each step. In this case, the number of protests per year and per period are calculated. 

### NI Shapefile 

```{r}
NI <- read_sf("data/NI.shp")%>% 
  dplyr::select(NAME_3, geometry) %>% 
  rename(District = NAME_3) %>% 
  mutate(
    District = dplyr::recode(District, "Newry and Mourne" = "Newry & Mourne"))
```

### Protests by year

```{r message=FALSE, warning=FALSE}
df <- read_csv("Databases/Final_Derry.csv")

data_year <- df |> 
  dplyr::select(-...1) %>% 
  group_by(year = floor_date(week, unit = "year"), District) %>% 
  summarise(
    yearly_protests = sum(n_protests)) |> 
  ungroup()|> 
  mutate(
    year = year(year))

pyear <- NI |> 
  left_join(data_year, by = "District")
  
```

### Protest by period

```{r}
data1980 <- data_year |> 
  filter(year <= 1983) |> 
  group_by(District) |> 
  summarise(n_protests = sum(yearly_protests)) |> 
  mutate(Period = "1979-83")

data1984 <- data_year |> 
  filter(year > 1983 & year <= 1987) |> 
  group_by(District) |> 
  summarise(n_protests = sum(yearly_protests)) |> 
  mutate(Period = "1984-1987")

data1988 <- data_year |> 
  filter(year > 1987 &year <= 1991) |> 
  group_by(District) |> 
  summarise(n_protests = sum(yearly_protests)) |> 
  mutate(Period = "1988-91")

data1991 <- data_year |> 
  filter(year > 1991) |> 
  group_by(District) |> 
  summarise(n_protests = sum(yearly_protests)) |> 
  mutate(Period = "1992-1995")

period_data <- rbind(data1980, data1984, data1988, data1991) |> 
  subset(District != "Other") 

pperiod <- NI |> 
  left_join(period_data, by = "District") 
```

## The Maps

Three main maps are constructed. 

### Fonts 

```{r}
sysfonts::font_add_google("Gideon Roman", family = "gideon roman")
showtext::showtext_auto()  
```

### Protests per year, static. 

The first one involve the representations of the log number of protests per district, faceting per each year. 

```{r}
pm <- ggplot()+
  geom_sf(data = pyear, aes(fill = log(yearly_protests + 1)))+
  scale_fill_gradient(low = "lightyellow", high = "red")+
  theme_map()+
  labs(title = "Number of protests per year")+
  theme(
    plot.title = element_text(face = "bold", 
                              size = 10, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    legend.position = "none") +
  facet_wrap(~year)
```

```{r}
pm
```

### Protests per year, animated. 

The second map is the animated version of the first representation. A separate markdown can be found with more examples. 

```{r message=FALSE, warning=FALSE}
ma <- ggplot()+
  geom_sf(data = pyear, aes(fill = log(yearly_protests +1 )))+
  geom_sf_label(data = pyear, aes(label = District), size = 2.5)+
  scale_fill_gradient(low = "lightyellow", high = "red")+ 
  annotation_north_arrow(location = "tr", 
                         which_north = "true",
                         style = north_arrow_fancy_orienteering, 
                         height = unit(2, "cm"), 
                         width = unit(2, "cm"))+
  annotation_scale(location = "bl") +
  theme_map()+
  transition_manual(year)+
  ggtitle("Year: {frame + 1978}", 
          subtitle = "Number of protests per year")+
  theme(
    plot.title = element_text(face = "bold", 
                              size = 14, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    plot.subtitle = element_text(size = 12, 
                                 hjust = 0.5, 
                                 family = "gideon roman"),
  legend.position = "none") +
  labs(x = NULL, y = NULL)
```

```{r message=FALSE, warning=FALSE}
pma <- animate(ma, fps = 5)
pma
```

### Protests by period (Annex I)

Third map: geographical representation of the log number of protests in four periods: 

1. 1979-1983
2. 1984-1987
3. 1988-1991
4. 1992-1995

```{r}
pm2 <- ggplot()+
  geom_sf(data = pperiod, aes(fill = log(n_protests + 1)))+
  scale_fill_gradient(low = "lightyellow", high = "red")+
  theme_map()+
  labs(title = "Number of protests per period")+
  theme(
    plot.title = element_text(face = "bold", 
                              size = 10, 
                              hjust = 0.5, 
                              family = "gideon roman"), 
    legend.position = "none") +
  facet_wrap(~Period)
```

```{r}
pm2
```

