---
title: "aggregating"
author: "Alejandro Aísa"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(rvest)
library(xml2)
library(readxl)
library(RcppRoll)
```

This is the second step of the thesis' data management. Assuming that the data is already harvested and cleaned, now I am able to format it and create the dependent variable at the aggregation level needed for tha analysis. 

## 1. Loading data 

First, I need to load the dataframes cleaned and filtered for the territory of Northern Ireland and the football data. 

```{r}
Protests <-  read_csv("Databases/Protests/ProtestsFiltered.csv") %>% 
  dplyr::select(-...1) %>% 
  filter(Country == "UK/Northern Ireland") 

Derry <- read_csv("Databases/DerryCity/DerryCityCleaned.csv") %>% 
  dplyr::select(-...1) %>% 
  filter(Competition == "Prem" | Competition == "First") %>% 
  group_by(week = floor_date(Date, unit = "week"), District) %>% 
  transmute(
    DerryHome, 
    DerryWin) %>% 
  ungroup()

Cliftonville <- read_csv("~/CliftonvilleCleaned.csv") %>% 
  dplyr::select(-...1) %>% 
  group_by(week = floor_date(MatchDate, unit = "week"), District) %>% 
  transmute(
    CliftonvilleHome,  
    CliftonvilleWin) %>% 
  ungroup()
```

## 2. Aggregating

First, I will create a sequence that accounts for all the days ranging from the first of January 1980 to the 31st of December 1995. i.e. the period of study. 

```{r}
date <- seq(as.Date("1980/01/01"), as.Date("1995/12/31"), by = "day")
```

Also, I obtain the number of protests per day. This step is necessary to perform the future aggregations. 

```{r}
Agg_protests <- Protests %>% 
  group_by(Date, District) %>% 
  summarise(
    n_protests = n(), 
    n_protesters = sum(N_Protesters)) %>%  
  ungroup()
```

### 2.1. Week-District Aggregation

At this step, I will create the dependent variable accounting for the number of protests per district. For that, I join the sequence with the date with the protests. Then, I just need to `complete` the dataframe with '0s'. It is necessary to establish that days without protests are not missing data. Finally, I just group by week and district and sum the number of protests that took place. Similarly, the binary independent variable is created. 

```{r}
weekly_protests <- as_tibble(date) %>%
  rename(Date = value) %>%
  left_join(Agg_protests, by = c("Date")) %>%
  complete(Date, District) %>% 
  drop_na(District) %>% 
  mutate(
    n_protests = coalesce(n_protests, 0), 
    n_protesters = coalesce(n_protesters, 0)) %>% 
  group_by(week = floor_date(Date, unit = "week"), District) %>% 
  summarise(
    n_protests = sum(n_protests), 
    n_protesters = sum(n_protesters)) %>% 
  mutate(
    is_protest = ifelse(n_protests >= 1, 1, 0)) 
```

#### 2.1.1. Derry aggregated 

With the data of protests aggregated by week, I can finally merge it with the Derry City FC results. 

```{r}
Derry_aggregated <-  weekly_protests  %>% 
  left_join(Derry, by = c("week", "District")) %>% 
  mutate(
    DerryHome = coalesce(DerryHome, 0), 
    DerryWin = coalesce(DerryWin, 0)) 
    
  
write.csv(Derry_aggregated, "Databases/Derry_aggregated.csv")
```

#### 2.1.2. Cliftonville aggregated 

Likewise, we can merge protests and Cliftonvile results. 

```{r}
Cliftonville_aggregated <- weekly_protests %>% 
  left_join(Cliftonville, by = c("week", "District")) %>% 
  mutate(
    CliftonvilleHome = coalesce(CliftonvilleHome, 0), 
    CliftonvilleWin = coalesce(CliftonvilleWin, 0)) 

write.csv(Cliftonville_aggregated, "Databases/Cliftonville_aggregated.csv")
```
## 3. Adding control variables 

When we have the dependent and independent variables together in a single dataframe, we may add the control variables. 

### 3.1. Loading the dataframes 

Luckily, I have been working in different dataframes for each particular variable. Thus, I may load the control variables and the (in)dependent variables at once:

```{r}
Religion <- read_csv("Databases/Religion.csv") %>% 
  dplyr::select(-...1)

Religion1981 <- Religion %>% 
  select(District, pcg1981)

Religion1991 <- Religion %>% 
  select(District, pcg1991)

Belfast_Distances <- read_csv("Databases/Belfast_Distances.csv")%>% 
  dplyr::select(-...1, -Distance)

Derry_Distances <- read_csv("Databases/Derry_Distances.csv")%>% 
  dplyr::select(-...1, -Distance)

control1981 <- read.csv("Databases/control1981.csv") %>% 
  dplyr::select(District, pcg_car, pcg_Unem)

control1991 <- read.csv("Databases/control1991.csv") %>% 
  dplyr::select(District, pcg_car, pcg_Unem) %>% 
  drop_na()

Cliftonville_aggregated <- read_csv("Databases/Cliftonville_aggregated.csv") %>%
  dplyr::select(-...1)

Derry_aggregated <- read_csv("Databases/Derry_aggregated.csv") %>% 
  dplyr::select(-...1)

```

### 3.2. Adding controls to the Derry Database

As stated in the thesis, for the period 1980-1987, the data from the 1981 census will be assigned. Then, the census of 1991 will provide the data for the period 1988-1995. 

Taking advantage of the workflow, I just need to `left_join` the different dataframes to have the final one. 

```{r}
DA_81_87 <- Derry_aggregated %>% 
  mutate(
    week = ymd(week)) %>% 
  filter(week < "1987-01-01") %>% 
  left_join(Religion1981, by = "District") %>% 
  rename(pcg_catholics = pcg1981) %>% 
  replace_na(list(pcg_catholics = 38.8, Catholic_majority = 0)) %>% 
  left_join(control1981, by = "District") %>% 
  replace_na(list(pcg_car = 0.81, pcg_Unem = 0.19)) 

DA_87_95 <- Derry_aggregated %>% 
  mutate(
    week = ymd(week)) %>% 
  filter(week >= "1987-01-01") %>% 
  left_join(Religion1991, by = "District") %>% 
  rename(pcg_catholics = pcg1991) %>% 
  replace_na(list(pcg_catholics = 38.8, Catholic_majority = 0)) %>% 
  left_join(control1991, by = "District") %>% 
  replace_na(list(pcg_car = 0.68, pcg_Unem = 0.16)) 

Final_Derry <- rbind(DA_81_87, DA_87_95)%>% 
  left_join(Derry_Distances, by = "District")

write.csv(Final_Derry, "Databases/Final_Derry.csv")
```

### 3.3. Adding control to the Cliftonville database

Same as before, I just need to merge the different databases:

```{r}
CA_81_87 <- Cliftonville_aggregated %>% 
  mutate(
    week = ymd(week)) %>% 
  filter(week < "1987-01-01") %>% 
  left_join(Religion1981, by = "District") %>% 
  rename(pcg_catholics = pcg1981) %>% 
  replace_na(list(pcg_catholics = 38.8, Catholic_majority = 0)) %>% 
  left_join(control1981, by = "District") %>% 
  replace_na(list(pcg_car = 0.81, pcg_Unem = 0.19)) 

CA_87_95 <- Cliftonville_aggregated %>% 
  mutate(
    week = ymd(week)) %>% 
  filter(week >= "1987-01-01") %>% 
  left_join(Religion1991, by = "District") %>% 
  rename(pcg_catholics = pcg1991) %>% 
  replace_na(list(pcg_catholics = 38.8, Catholic_majority = 0)) %>% 
  left_join(control1991, by = "District") %>% 
  replace_na(list(pcg_car = 0.68, pcg_Unem = 0.16)) 

Final_Cliftonville <- rbind(CA_81_87, CA_87_95) %>% 
  left_join(Belfast_Distances, by = "District")

write.csv(Final_Cliftonville, "Databases/Final_Cliftonville.csv")
```

The resulting dataframes are ready for the analysis. Now they contain the data aggregated at the week-district level, accounting for the number of protests, whether the teams played at home or won a match. The control variables accounting for the percentage of catholics, the distance to the City, and the socio-economic variables are also included. 